Trilemma

### 2023-06-14 Trilemma_clustering.Rmd
paper revisions (graphs)

dat <- ecofp %>%
  filter(record == "EFConsPerCap") %>%
  select(countryName, isoa2, year, value) %>% 
  rename(EFconsPerCap = value, country = countryName) %>% 
  filter(year >= 1995)

dat <- dat %>% 
  left_join(
    .,
    inq %>% 
      filter(!is.na(gini_reported), year >= 1995) %>% 
      select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
      group_by(country, year, isoa2, country_code) %>% 
      summarize(gini_mean = mean(gini_reported, na.rm = TRUE))

@@

dat <- dat %>% 
  left_join(
    ., 
    dat1 %>% 
      select(-last_col()) %>% 
      filter(indicator_name == "GNI per capita (US$)") %>% 
      pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>%
      mutate(year = str_remove(year, "x")) %>% 
      mutate(year = as.numeric(year)) %>% 
      mutate(ismissing = is.na(value)) %>% 
      filter(year >= 1995) %>% 
      rename(country = country_name, gni = value)
  )

@@

Compute missing values for inequality:
```{r}
missingness <- dat %>% 
    mutate(ismissing = is.na(gini_mean)) %>% 
    group_by(country) %>% 
    summarize(missingness = sum(ismissing)/n()) %>% 
    arrange(missingness)

key_countries <- missingness %>% 
    filter(missingness <= 0.3) %>% 
    pull(country)

droped_countries <- missingness %>% 
    filter(missingness > 0.3) %>% 
    pull(country)
```

```{r}
#library(tsibbledata)
df_dat <- dat %>% 
    select(-starts_with("indicator")) %>% 
    filter(country %in% key_countries) %>% 
    as_tsibble(key = country, index = year)

@@

df_dat <- df_dat %>% 
    group_by(country) %>% 
    mutate(
        gini = na_interpolation(gini_mean),
        gni2 = na_interpolation(gni),
        EFconsPerCap = na_interpolation(EFconsPerCap)
    ) 
```

@@

```{r}
df_dat %>% 
    pivot_longer(cols = c("EFconsPerCap", "gini", "gni2"),
                 names_to = "variable", values_to = "value") %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(group = country)) +
    facet_wrap(~variable, ncol = 1, scales = "free") +
    theme_light(base_size = 15)
```

```{r}
p1 <- df_dat %>% 
    ggplot(aes(EFconsPerCap, gini, group = country)) +
    geom_point(size = 0.5) +
    geom_path(size = 0.2, aes(color = year)) +
    annotate(geom = "rect", xmin = 0, xmax = 1.68, ymin = 20, ymax = 30,
             fill = "skyblue", alpha = 0.2) +
    annotate(geom ="rect",xmin = 0, xmax = 1.68, ymin = 30, ymax = 62,
             fill = "orange", alpha = 0.2) +
    annotate(geom ="rect", xmin = 1.68, xmax = 18, ymin = 20, ymax = 30,
             fill = "orange", alpha = 0.2) +
    annotate(geom="rect", xmin = 1.68, xmax = 18, ymin = 30, ymax = 62,
             fill = "red", alpha = 0.2) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
    labs(y = "Gini", x = "Ecological footprint", tag = "A") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

p2<- df_dat %>% 
  ggplot(aes(gni2, gini, group = country)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  geom_path(size = 0.2, aes(color = year)) +
    annotate(geom = "rect", xmin = 10, xmax = 12746, ymin = 20, ymax = 30,
             fill = "orange", alpha = 0.2) +
    annotate(geom ="rect",xmin = 10, xmax = 12746, ymin = 30, ymax = 62,
             fill = "red", alpha = 0.2) +
    annotate(geom ="rect", xmin = 12746, xmax = 105000, ymin = 20, ymax = 30,
             fill = "skyblue", alpha = 0.2) +
    annotate(geom="rect", xmin = 12746, xmax = 105000, ymin = 30, ymax = 62,
             fill = "orange", alpha = 0.2) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 12746, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
      scale_x_log10() +
    labs(y = "Gini", x = "GNI [log]", tag = "B") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

p3<- df_dat %>% 
  ggplot(aes(gni2, EFconsPerCap, group = country)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  geom_path(size = 0.2, aes(color = year))+
    annotate(geom = "rect", xmin = 10, xmax = 12746, ymin = 0, ymax = 1.68,
             fill = "orange", alpha = 0.2) +
    annotate(geom ="rect",xmin = 10, xmax = 12746, ymin = 1.68, ymax = 18,
             fill = "red", alpha = 0.2) +
    annotate(geom ="rect", xmin = 12746, xmax = 105000, ymin = 0, ymax = 1.68,
             fill = "skyblue", alpha = 0.2) +
    annotate(geom="rect", xmin = 12746, xmax = 105000, ymin = 1.68, ymax = 18,
             fill = "orange", alpha = 0.2) +
    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 12746, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
      scale_x_log10() +
    labs(y = "Ecological footprint", x = "GNI [log]", tag = "C") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

@@

p1 + p2 + p3 + 
    plot_layout(guides = "collect") & theme_light(base_size = 18) & 
    theme(legend.position = "bottom", panel.grid = element_blank(),
          legend.key.width = unit(20, "mm"), legend.key.height = unit(3,"mm"))
```

```{r}
df_dat %>% 
    as_tibble() %>% 
    group_by(country) %>% 
    summarize(
        mean_ef = mean(EFconsPerCap),
        mean_gni = mean(gni2),
        mean_gini = mean(gini),
        var_ef = var(EFconsPerCap),
        var_gni = var(gni2),
        var_gini = var(gini)
    ) %>% 
    arrange(desc(mean_ef))
```
### Experimenting with phase plane

@@

```{r}
df_dat |> 
    as_tibble() |> 
    ungroup() |> 
    group_by(country) |> 
    summarize(mgt = sum(magni, na.rm = TRUE), 
              theta = sum(theta, na.rm = TRUE),
              gamma = sum(gamma, na.rm = TRUE)) |> 
    arrange(desc(mgt))
```
```{r fig.width=6, fig.height=3}
p <- df_dat |> 
    as_tibble() |> 
    select(country, year, d_EF, d_gini, d_gni) |> 
    pivot_longer(cols = starts_with("d_"), names_to = "vars", values_to = "vals") |> 
    filter(!is.na(vals)) |> 
    mutate(vars = case_when(
        vars == "d_EF" ~ "Ecological Footprint",
        vars == "d_gni" ~ "GNI [log]",
        vars == "d_gini" ~ "Gini"
    )) |> 
    ggplot(aes(year, vals, group = country)) +
    geom_path(size = 0.2, alpha = 0.5) +
    facet_wrap(~vars, scales = "free") +
    labs(x = "Year", y = "First difference") +
    theme_light(base_size = 7)

# ggsave(
#     plot = p, path = "figures/", file = "first_diff.png", device = "png",
#     width = 6, height = 3, dpi = 300, bg = "white")
p + theme_light(base_size = 20)
```
```{r fig.width=3, fig.height=3}
c1 <- df_dat |> 
    ggplot(aes(EFconsPerCap, gini)) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
    geom_density2d(size = 0.25, color = "#446B9C") +
    geom_point(aes(color = gni_log), size = 0.5) +
    geom_path(aes(group = country), size = 0.1) +
    scico::scale_color_scico("Prosperity (GNI [log])") +
    labs(y = "Inequality (Gini)", x = "Ecological footprint") +
    theme_light(base_size = 6) +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(7, "mm"), legend.key.height = unit(2,"mm"))

c1 + theme_light(base_size = 18)
```
```{r fig.width=3, fig.height=3}
c2 <- df_dat |> 
    ggplot(aes(gni2, gini)) +
    geom_density2d(size = 0.25, color = "#446B9C") +
    geom_point(aes(color = EFconsPerCap), size = 0.5) +
    geom_path(aes(group = country), size = 0.1) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
    scico::scale_color_scico("Ecological footprint",palette = "berlin") +
    labs(x= "Prosperity (GNI [log])", y = "Inequality (Gini)") +
    scale_x_log10() +
    theme_light(base_size = 6) +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(7, "mm"), legend.key.height = unit(3,"mm"))

c2 + theme_light(base_size = 7)
```
```{r fig.width=3, fig.height=3}
c3 <- df_dat |> 
    ggplot(aes(gni2, EFconsPerCap)) +
    geom_density2d(size = 0.25, color = "#446B9C") +
    geom_point(aes(color = gini), size =0.5) +
    geom_path(aes(group = country), size = 0.1) +
    scico::scale_color_scico("Inequality (Gini)") +
    labs(x = "Prosperity (GNI)", y = "Ecological footprint") +
    scale_x_log10()+
    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50") +
    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
    theme_light(base_size = 6) +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(7, "mm"), legend.key.height = unit(3,"mm"))

c3 + theme_light(base_size = 18)
```
Who is the weird trajectory out there, world? No it's Luxemburg

@@

#save(df_dat, file = "data/cleaned_trilemma.RData")
```

@@

```{r fig.width=6, fig.height=2}
df_summary <- df_dat |> 
    as_tibble() |> 
    ungroup() |> 
    group_by(country) |> 
    dplyr::summarize(sd_magnitude = sd(magni, na.rm = TRUE),
              sd_theta = sd(theta, na.rm = TRUE),
              sd_gamma = sd(gamma, na.rm = TRUE), 
              EF = mean(EFconsPerCap, na.rm = TRUE),
              Gini = mean(gini, na.rm = TRUE),
              GNI = mean(gni_log, na.rm = TRUE),
              sd_EF = sd(EFconsPerCap, na.rm = TRUE),
              sd_Gini = sd(gini, na.rm = TRUE),
              sd_GNI = sd(gni_log, na.rm = TRUE),
              #magnitude_m = mean(magni, na.rm = TRUE),
              #theta_m = mean(theta, na.rm = TRUE),
              #gamma_m = mean(gamma, na.rm = TRUE),
              magnitude = sum(magni, na.rm = TRUE),
              theta = sum(theta, na.rm = TRUE),
              gamma = sum(gamma, na.rm = TRUE)
              )

@@

library(clValid)
library(mclust)
## Stability & Internal validation
stab <- clValid::clValid(
    obj = df_summary %>% as.data.frame() %>% 
        select(where(is.numeric)) %>%
        as.matrix(),
    nClust = 3,
    clMethods = c(
        "hierarchical", "kmeans",  "som", 
        "model", "diana", "sota", "pam", "clara", "agnes"),
    validation = c('stability', "internal"),
    #metric = "manhattan",
    method = "ward",
    verbose = FALSE
) ## Hierarchical is the optimal method

@@

```{r fig.width=7, fig.height=8}
df_summary |> 
    mutate(clus = stab@clusterObjs$kmeans$`3`$cluster ) |> 
    mutate(country = as_factor(country)) |> 
    mutate(country = fct_reorder2(country, magnitude, Gini)) |> 
    pivot_longer(
        cols = sd_magnitude:gamma, names_to = "vars", 
        values_to = "vals") |> 
    ggplot(aes(vals, country)) +
    geom_col(aes(fill = as.factor(clus))) +
    facet_grid(clus~vars, scales = "free") +
    theme_light(base_size = 10)
```

```{r message = FALSE, warning = FALSE, fig.width=5, fig.height=5}
df_summary |>
    select(-starts_with("country")) |> 
    GGally::ggpairs()
```

>>>>

```{r}
## there are different country name spellings, so work only with iso codes and recover names later if you need them.
countries <- read_csv(
    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
    janitor::clean_names()

dat <- ecofp %>%
  filter(record == "EFConsPerCap") %>%
  select(countryName, isoa2, year, value) %>% 
  rename(EFconsPerCap = value, country = countryName) %>% 
    filter(!is.na(isoa2)) |> 
  filter(year >= 1995)

## correct Sudan:
dat$isoa2[dat$country == "Sudan (former)"] <- "SD"
## get rid of country names
dat <- dat |> select(-country)

## add iso3 code
dat <- dat |> 
    left_join(countries |> select(iso3 = country_code, isoa2 = x2_alpha_code))

dat$iso3[dat$isoa2 == "NA"] <- "NAM" # correct Namibia

dat <- dat %>% 
  left_join(
    .,
    inq %>% 
      filter(!is.na(gini_reported), year >= 1995) %>% 
      select(country, isoa2 = c2, iso3 = c3, year, gini_reported) %>% 
      group_by(country, year, isoa2, iso3) %>% 
      summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) |> 
        ungroup() |> select(-country) # remove country names

@@

dat <- dat %>% 
  left_join(
    ., 
    dat1 %>% 
      select(-last_col()) %>% 
      filter(indicator_name == "GNI per capita (US$)") %>% 
      pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>%
      mutate(year = str_remove(year, "x")) %>% 
      mutate(year = as.numeric(year)) %>% 
      mutate(ismissing = is.na(value)) %>% 
      filter(year >= 1995) %>% 
      rename(country = country_name, gni = value, iso3 = country_code) |> 
        ungroup() |> select(-country)
  )

@@

Compute missing values for inequality:
```{r}
missingness <- dat %>% 
    mutate(ismissing = is.na(gini_mean)) %>% 
    group_by(iso3) %>% 
    summarize(missingness = sum(ismissing)/n()) %>% 
    arrange(missingness)

key_countries <- missingness %>% 
    filter(missingness <= 0.3) %>% 
    pull(iso3)

droped_countries <- missingness %>% 
    filter(missingness > 0.3) %>% 
    pull(iso3)
```

```{r}
#library(tsibbledata)
df_dat <- dat %>% 
    select(-starts_with("indicator")) %>% 
    filter(iso3 %in% key_countries) %>% 
    as_tsibble(key = iso3, index = year)

@@

df_dat <- df_dat %>% 
    group_by(iso3) %>% 
    mutate(
        gini = na_interpolation(gini_mean),
        gni2 = na_interpolation(gni),
        EFconsPerCap = na_interpolation(EFconsPerCap)
    ) 
```

@@

Use this graph for trajectories of all countries later

```{r}
df_dat %>% 
    pivot_longer(cols = c("EFconsPerCap", "gini", "gni2"),
                 names_to = "variable", values_to = "value") %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(group = iso3)) +
    facet_wrap(~variable, ncol = 1, scales = "free") +
    theme_light(base_size = 15)
```

```{r}
p1 <- df_dat %>% 
    ggplot(aes(EFconsPerCap, gini, group = iso3)) +
    geom_point(size = 0.5, aes(color = year)) +
    geom_path(size = 0.2, aes(color = year)) +
    annotate(geom = "rect", xmin = 0, xmax = 1.68, ymin = 20, ymax = 30,
             fill = "skyblue", alpha = 0.2) +
    annotate(geom ="rect",xmin = 0, xmax = 1.68, ymin = 30, ymax = 62,
             fill = "orange", alpha = 0.2) +
    annotate(geom ="rect", xmin = 1.68, xmax = 18, ymin = 20, ymax = 30,
             fill = "orange", alpha = 0.2) +
    annotate(geom="rect", xmin = 1.68, xmax = 18, ymin = 30, ymax = 62,
             fill = "red", alpha = 0.2) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
    labs(y = "Gini", x = "Ecological footprint", tag = "A") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

p2<- df_dat %>% 
  ggplot(aes(gni2, gini, group = iso3)) +
  geom_point(size = 0.5, show.legend = FALSE, aes(color = year)) +
  geom_path(size = 0.2, aes(color = year)) +
    annotate(geom = "rect", xmin = 10, xmax = 12746, ymin = 20, ymax = 30,
             fill = "orange", alpha = 0.2) +
    annotate(geom ="rect",xmin = 10, xmax = 12746, ymin = 30, ymax = 62,
             fill = "red", alpha = 0.2) +
    annotate(geom ="rect", xmin = 12746, xmax = 105000, ymin = 20, ymax = 30,
             fill = "skyblue", alpha = 0.2) +
    annotate(geom="rect", xmin = 12746, xmax = 105000, ymin = 30, ymax = 62,
             fill = "orange", alpha = 0.2) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 12746, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
      scale_x_log10() +
    labs(y = "Gini", x = "GNI [log]", tag = "B") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

p3<- df_dat %>% 
  ggplot(aes(gni2, EFconsPerCap, group = iso3)) +
  geom_point(size = 0.5, show.legend = FALSE, aes(color = year)) +
  geom_path(size = 0.2, aes(color = year))+
    annotate(geom = "rect", xmin = 10, xmax = 12746, ymin = 0, ymax = 1.68,
             fill = "orange", alpha = 0.2) +
    annotate(geom ="rect",xmin = 10, xmax = 12746, ymin = 1.68, ymax = 18,
             fill = "red", alpha = 0.2) +
    annotate(geom ="rect", xmin = 12746, xmax = 105000, ymin = 0, ymax = 1.68,
             fill = "skyblue", alpha = 0.2) +
    annotate(geom="rect", xmin = 12746, xmax = 105000, ymin = 1.68, ymax = 18,
             fill = "orange", alpha = 0.2) +
    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 12746, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
      scale_x_log10() +
    labs(y = "Ecological footprint", x = "GNI [log]", tag = "C") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

@@

p1 + p2 + p3 + 
    plot_layout(guides = "collect") & theme_light(base_size = 12) & 
    theme(legend.position = "bottom", panel.grid = element_blank(),
          legend.key.width = unit(20, "mm"), legend.key.height = unit(3,"mm"))
```

```{r}
df_dat %>% 
    as_tibble() %>% 
    group_by(iso3) %>% 
    summarize(
        mean_ef = mean(EFconsPerCap),
        mean_gni = mean(gni2),
        mean_gini = mean(gini),
        var_ef = var(EFconsPerCap),
        var_gni = var(gni2),
        var_gini = var(gini)
    ) %>% 
    arrange(desc(mean_ef))
```
### Experimenting with phase plane

@@

```{r}
df_dat |> 
    as_tibble() |> 
    ungroup() |> 
    group_by(iso3) |> 
    summarize(mgt = sum(magni, na.rm = TRUE), 
              theta = sum(theta, na.rm = TRUE),
              gamma = sum(gamma, na.rm = TRUE)) |> 
    arrange(desc(mgt))
```
```{r fig.width=6, fig.height=3}
p <- df_dat |> 
    as_tibble() |> 
    select(iso3, year, d_EF, d_gini, d_gni) |> 
    pivot_longer(cols = starts_with("d_"), names_to = "vars", values_to = "vals") |> 
    filter(!is.na(vals)) |> 
    mutate(vars = case_when(
        vars == "d_EF" ~ "Ecological Footprint",
        vars == "d_gni" ~ "GNI [log]",
        vars == "d_gini" ~ "Gini"
    )) |> 
    ggplot(aes(year, vals, group = iso3)) +
    geom_path(size = 0.2, alpha = 0.5) +
    facet_wrap(~vars, scales = "free") +
    labs(x = "Year", y = "First difference") +
    theme_light(base_size = 7)

# ggsave(
#     plot = p, path = "figures/", file = "first_diff.png", device = "png",
#     width = 6, height = 3, dpi = 300, bg = "white")
p + theme_light(base_size = 20)
```
```{r fig.width=3, fig.height=3}
c1 <- df_dat |> 
    ggplot(aes(EFconsPerCap, gini)) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
    geom_density2d(size = 0.25, color = "#446B9C") +
    geom_point(aes(color = gni_log), size = 0.5) +
    geom_path(aes(group = iso3), size = 0.1) +
    scico::scale_color_scico("Prosperity (GNI [log])") +
    labs(y = "Inequality (Gini)", x = "Ecological footprint") +
    theme_light(base_size = 6) +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(7, "mm"), legend.key.height = unit(2,"mm"))

c1 
```
```{r fig.width=3, fig.height=3}
c2 <- df_dat |> 
    ggplot(aes(gni2, gini)) +
    geom_density2d(size = 0.25, color = "#446B9C") +
    geom_point(aes(color = EFconsPerCap), size = 0.5) +
    geom_path(aes(group = iso3), size = 0.1) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
    scico::scale_color_scico("Ecological footprint",palette = "berlin") +
    labs(x= "Prosperity (GNI [log])", y = "Inequality (Gini)") +
    scale_x_log10() +
    theme_light(base_size = 6) +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(7, "mm"), legend.key.height = unit(3,"mm"))

c2 + theme_light(base_size = 7)
```
```{r fig.width=3, fig.height=3}
c3 <- df_dat |> 
    ggplot(aes(gni2, EFconsPerCap)) +
    geom_density2d(size = 0.25, color = "#446B9C") +
    geom_point(aes(color = gini), size =0.5) +
    geom_path(aes(group = iso3), size = 0.1) +
    scico::scale_color_scico("Inequality (Gini)") +
    labs(x = "Prosperity (GNI)", y = "Ecological footprint") +
    scale_x_log10()+
    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50") +
    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
    theme_light(base_size = 6) +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(7, "mm"), legend.key.height = unit(3,"mm"))

c3 
```
Who is the weird trajectory out there, world? No it's Luxemburg

@@

# save(df_dat, file = "data/cleaned_trilemma-20230614.RData")
```

@@

```{r fig.width=6, fig.height=4}
df_summary <- df_dat |> 
    as_tibble() |> 
    ungroup() |> 
    group_by(iso3) |> 
    dplyr::summarize(sd_magnitude = sd(magni, na.rm = TRUE),
              sd_theta = sd(theta, na.rm = TRUE),
              sd_gamma = sd(gamma, na.rm = TRUE), 
              EF = mean(EFconsPerCap, na.rm = TRUE),
              Gini = mean(gini, na.rm = TRUE),
              GNI = mean(gni_log, na.rm = TRUE),
              sd_EF = sd(EFconsPerCap, na.rm = TRUE),
              sd_Gini = sd(gini, na.rm = TRUE),
              sd_GNI = sd(gni_log, na.rm = TRUE),
              #magnitude_m = mean(magni, na.rm = TRUE),
              #theta_m = mean(theta, na.rm = TRUE),
              #gamma_m = mean(gamma, na.rm = TRUE),
              magnitude = sum(magni, na.rm = TRUE),
              theta = sum(theta, na.rm = TRUE),
              gamma = sum(gamma, na.rm = TRUE)
              )

@@

library(clValid)
library(mclust)
## Stability & Internal validation
stab <- clValid::clValid(
    obj = df_summary %>% as.data.frame() %>% 
        select(where(is.numeric)) %>%
        as.matrix(),
    nClust = 2:9,
    clMethods = c(
        "hierarchical", "kmeans",  "som", 
        "model", "diana", "sota", "pam", "clara", "agnes"),
    validation = c('stability', "internal"),
    #metric = "manhattan",
    method = "ward",
    verbose = FALSE
) ## Hierarchical is the optimal method

@@

```{r fig.width=7, fig.height=8}
df_summary |> 
    mutate(clus = stab@clusterObjs$kmeans$`3`$cluster ) |> 
    mutate(iso3 = as_factor(iso3)) |> 
    mutate(iso3 = fct_reorder2(iso3, magnitude, Gini)) |> 
    pivot_longer(
        cols = sd_magnitude:gamma, names_to = "vars", 
        values_to = "vals") |> 
    ggplot(aes(vals, iso3)) +
    geom_col(aes(fill = as.factor(clus))) +
    facet_grid(clus~vars, scales = "free") +
    theme_light(base_size = 10)
```

```{r message = FALSE, warning = FALSE, fig.width=5, fig.height=5}
df_summary |>
    select(-starts_with("iso3")) |> 
    GGally::ggpairs()
```
